Forecasting the Support of Kamala Harris in 2024 U.S. Presidential Election using a Model-Based Approach with Poll-of-Polls Data

### 2024-11-04 paper/ForecastingtheSupportofKamalaHarrisin2024U.S.PresidentialElection.qmd
update paper caption

### Recency {#sec-recency}
```{r}
#| label: fig-recency
#| fig-cap: Distribution of the Poll Recency since July 21, 2024
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = recency)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Recency (days)",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r}
#| label: fig-numericgrade
#| fig-cap: Distribution of the Numeric Grade of the Polls
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-percentage
#| fig-cap: Distribution of the Support Percentage of Kamala Harris
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = pct)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Support for Harris (%)",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r}
#| label: tbl-category
#| tbl-cap: Summary Statistics of the categorical variables population and national_pull
#| echo: false
#| warning: false
#| message: false
harris_data|>select(population,national_poll)|>
  datasummary_skim(type = "categorical")
```
### National Poll {#sec-national-poll}

@@

```{r}
#| label: fig-state
#| fig-cap: Count of Polls by the US state 
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-pollster
#| fig-cap: Frequency of Polls by Pollster
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-pollstertime
#| fig-cap: Harris Support Over Time by Pollster (Top 6 Pollsters)
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-numericgradetime
#| fig-cap: Harris Support Over Time by Numeric Grade of the Polls
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Harris Support Based on Polling Data"
#| label: tbl-summary1
msummary(
  list("Unweighted" = harris_unweighted_model, "Weighted" = harris_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska, the reference level of population is likely voter, and the reference level of pollsters is Angus Reid.")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Trump Support Based on Polling Data"
#| label: tbl-summary2
msummary(
  list("Unweighted" = trump_unweighted_model, "Weighted" = trump_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska,the reference level of population is likely voter, and the reference level of pollsters is ABC/Washington Post.")
```

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Harris Kamala leads Donald Trump an average 1% in Predicted Support in the 2024 US Election (From August 1, 2024, to November 2, 2024)"
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| label: fig-support-trend
# Plotting with the filtered data
plan(multisession)
##unweighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), method = "lm",size = 1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted),size = 0.2)+
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) +
  annotate("text", x = max(combined_predictions_filtered$end_date) -0.5, y = 48, 
           label = "47.5%", color = "blue", hjust = -0.2, size = 2) +  # Adjust size here
  annotate("text", x = max(combined_predictions_filtered$end_date) - 0.5, y = 46, 
           label = "46.5%", color = "red", hjust = -0.2, size = 2)

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, Michigan, and Wisconsin by Model Prediction"
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)
late<-combined_predictions_filtered[which.max(combined_predictions_filtered$end_date), ]
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate))+
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 0.5)+
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1, size =6),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| fig-cap: "Major Pollsters Siena/NYT and Quinnipiac favor Harris, while Emerson, AtlasIntel, and Beacon/Shaw present a more competitive picture"
#| label: fig-support-pollster
plan(multisession)
# Calculate the top 10 pollsters by frequency in the data
top_pollsters <- combined_data_total %>%
  count(pollster) %>%
  top_n(5, n) %>%
  pull(pollster)

>>>>

### Recency {#sec-recency}
```{r}
#| label: fig-recency
#| fig-cap: The distribution of the Poll Recency since July 21, 2024 highlights a strong focus on recent polling data, with the highest frequency of polls occurring within the first few days.
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = recency)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Recency (days)",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r}
#| label: fig-numericgrade
#| fig-cap: The distribution of the Numeric Grade of the Polls shows that surveys are conducted by highly rated pollsters, which ensure the reliability of the analysis.
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-percentage
#| fig-cap: The distribution of the Support Percentage of Kamala Harris clustered around 50%, with the majority of support falling between 40% and 55%.
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = pct)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Support for Harris (%)",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r}
#| label: tbl-category
#| tbl-cap: Summary statistics of the categorical variables population and national_pull reveal that the majority of the polls are state-level (68.3%) and focus primarily on likely voters (75%).
#| echo: false
#| warning: false
#| message: false
harris_data|>select(population,national_poll)|>
  datasummary_skim(type = "categorical")
```
### National Poll {#sec-national-poll}

@@

```{r}
#| label: fig-state
#| fig-cap: Count of Polls by the US state shows that Polling Distribution Prioritizes National Sentiment and High-Impact Battleground States.
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-pollster
#| fig-cap: Frequency of Polls Highlights Siena/NYT as Leading Pollster, Followed by YouGov and Emerson.
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-pollstertime
#| fig-cap: Trend Analysis of Kamala Harris Support from August to October Shows Upward Trends by Emerson and Beacon/Shaw, with Siena/NYT Showing Decline.
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-numericgradetime
#| fig-cap: Harris's Support Trends Stable Around 50%, with Minor Fluctuations Across Pollster Numeric Grades, Especially in Polls with Numeric Grade 2.5.
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Harris Support Based on Polling Data since 21 July 2024 (Harris first declared) to 2 November 2024"
#| label: tbl-summary1
msummary(
  list("Unweighted" = harris_unweighted_model, "Weighted" = harris_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska, the reference level of population is likely voter, and the reference level of pollsters is Angus Reid.")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Trump Support Based on Polling Data since 21 July 2024 (Harris first declared) to 2 November 2024"
#| label: tbl-summary2
msummary(
  list("Unweighted" = trump_unweighted_model, "Weighted" = trump_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska,the reference level of population is likely voter, and the reference level of pollsters is ABC/Washington Post.")
```

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Harris Kamala leads Donald Trump an average 1% in Predicted Support in the 2024 US Election (From August 1, 2024, to November 2, 2024)."
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| label: fig-support-trend
# Plotting with the filtered data
plan(multisession)
##unweighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.1)+ 
  geom_smooth(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), method = "lm",size = 1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted),size = 0.2)+
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) +
  annotate("text", x = max(combined_predictions_filtered$end_date) -0.5, y = 48, 
           label = "47.5%", color = "blue", hjust = -0.2, size = 2) +  # Adjust size here
  annotate("text", x = max(combined_predictions_filtered$end_date) - 0.5, y = 46, 
           label = "46.5%", color = "red", hjust = -0.2, size = 2)

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| fig-cap: "Harris is likely to lead Trump in Pennsylvania, Michigan, and Wisconsin by Model Prediction (From August 1, 2024, to November 2, 2024)."
#| label: fig-support-state
## unweighted
plan(multisession)
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)
late<-combined_predictions_filtered[which.max(combined_predictions_filtered$end_date), ]
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate))+
  geom_point(alpha = 0.1)+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 0.5)+
  facet_wrap(~ state,ncol = 4) +
  labs(
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Model: Candidate Support Percentage = national_poll + pollster + population + state") +
  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1, size =6),
        legend.box = "vertical",
        strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-subcap: ["Unweighted","Weighted"]
#| layout-ncol: 1
#| fig-cap: "Major Pollsters Siena/NYT and Quinnipiac favor Harris, while Emerson, AtlasIntel, and Beacon/Shaw present a more competitive picture."
#| label: fig-support-pollster
plan(multisession)
# Calculate the top 10 pollsters by frequency in the data
top_pollsters <- combined_data_total %>%
  count(pollster) %>%
  top_n(5, n) %>%
  pull(pollster)
